---
title: "Univariate EDA"
author: "Dynasty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Univariate EDA**

# Univariate Graphical Exploratory Data Analysis

## 1. Measures of Central Tendency

Before embarking on developing statistical models and generating predictions, it is essential to understand our data. This is typically done using conventional numerical and graphical methods.


We will be using the hills dataset, this dataset contains information on hill climbs made by various athletes

```{r}
library(MASS)
head(hills)
```


Rows and columns 

```{r}
dim(hills)
```

#### Mean Code Example 1.1 

Find the mean of the distance covered by the athletes and assigning the mean to the variable athletes.dist.mean

```{r}
athletes.dist.mean <- mean(hills$dist)
athletes.dist.mean
```

The mean distance covered is 7.528571


#### Median Code Example 1.2

Finding the median which is the middle most value of the distance covered dist

```{r}
athletes.dist.median <- median(hills$dist)
athletes.dist.median
```

The meadian is 6

#### Mode Code Example 1.3

Find the mode which is the value that has highest number of occurrences in a set of data. 

Unfotunately, R does not have a standard in-built function to calculate mode so we have to build one, We create the mode function that will perform our mode operation for us

```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```


Now we Calculate the mode using out getmode() function

```{r}
athletes.dist.mode <- getmode(hills$dist)
athletes.dist.mode
```

##### Let's Challenge Ourselves
Will Find the mean, median, mode of the total evening calls given the following dataset 

Dataset url = http://bit.ly/CustomerSignatureforChurnAnalysis

```{r}
library(data.table)
churn <- fread('http://bit.ly/CustomerSignatureforChurnAnalysis')
head(churn)
```

