---
title: "Univariate EDA"
author: "Dynasty"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Univariate EDA**

# Univariate Graphical Exploratory Data Analysis

## 1. Measures of Central Tendency

Before embarking on developing statistical models and generating predictions, it is essential to understand our data. This is typically done using conventional numerical and graphical methods.


We will be using the hills dataset, this dataset contains information on hill climbs made by various athletes

```{r}
library(MASS)
head(hills)
```


Rows and columns 

```{r}
dim(hills)
```

#### Mean Code Example 1.1 

Find the mean of the distance covered by the athletes and assigning the mean to the variable athletes.dist.mean

```{r}
athletes.dist.mean <- mean(hills$dist)
athletes.dist.mean
```

The mean distance covered is 7.528571


#### Median Code Example 1.2

Finding the median which is the middle most value of the distance covered dist

```{r}
athletes.dist.median <- median(hills$dist)
athletes.dist.median
```

The meadian is 6

#### Mode Code Example 1.3

Find the mode which is the value that has highest number of occurrences in a set of data. 

Unfotunately, R does not have a standard in-built function to calculate mode so we have to build one, We create the mode function that will perform our mode operation for us

```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```


Now we Calculate the mode using out getmode() function

```{r}
athletes.dist.mode <- getmode(hills$dist)
athletes.dist.mode
```

##### Let's Challenge Ourselves
Will Find the mean, median, mode of the total evening calls given the following dataset 

Dataset url = http://bit.ly/CustomerSignatureforChurnAnalysis

```{r}
library(data.table)
churn <- fread('http://bit.ly/CustomerSignatureforChurnAnalysis')
head(churn)
```



Let's see number of rows and columns

```{r}
dim(churn)
```

#### **Data cleaning**

Let's do some data cleaning


**Checking for Missing values**

```{r}
is.null(churn)
```

We don't have null values.



**Checking for Duplicates**


```{r}
churn_duplicated <- churn[duplicated(churn),]
churn_duplicated
```

```{r}
unique(churn)
```




We have no duplicates


**Checking for Outliers**
```{r}
library("ggplot2")                                           # Load ggplot2
```


We will focus on the numeric columns

```{r}
boxplot(churn$area_code)
```

```{r}
boxplot(churn$account_length)
```

```{r}
boxplot(churn$number_vmail_messages)
```

```{r}
boxplot(churn$total_day_minutes)
```

```{r}
boxplot(churn$total_day_calls)
```

```{r}
boxplot(churn$total_day_charge)
```

```{r}
boxplot(churn$total_eve_minutes)
```

```{r}
boxplot(churn$total_eve_calls)
```

```{r}
boxplot(churn$total_eve_charge)
```

```{r}
boxplot(churn$total_night_minutes)
```

```{r}
boxplot(churn$total_night_calls)
```

```{r}
boxplot(churn$total_night_charge)
```

```{r}
boxplot(churn$total_intl_minutes)
```

```{r}
boxplot(churn$total_intl_calls)
```

```{r}
boxplot(churn$total_intl_charge)
```


```{r}
boxplot(churn$number_customer_service_calls)
```








we have some columns with outliers


### **Questions**

**Find the minimum of total day calls**

```{r}
churn.dist.min <- min(churn$total_day_calls)
churn.dist.min
```

**Find the maximum i.e. max() total day calls**

```{r}
churn.dist.max <- max(churn$total_day_calls)
churn.dist.max
```

**Find the range i.e. range() of total day calls**
```{r}
churn.dist.range <- range(churn$total_day_calls)
churn.dist.range
```


**Find the quantile of total day calls**

```{r}
churn.dist.quantile <- quantile(churn$total_day_calls)
churn.dist.quantile
```


**Find the variance of total day calls**

```{r}
churn.dist.variance <- var(churn$total_day_calls)
churn.dist.variance
```


**Find the standard deviation of total day calls**
```{r}
churn.dist.sd <- sd(churn$total_day_calls)
churn.dist.sd
```




## **3. Univariate Graphical**

#### Box Plots Code Example 3.1


Lets create a boxplot graph for the distance using the boxplot() function

```{r}
boxplot(hills$dist)
```







The box plot of an observation variable is a graphical representation based on its quartiles, as well as its smallest and largest values. It attempts to provide a visual shape of the data distribution.

#### Bar Graph Code Example 3.2

A bar graph of a qualitative data sample consists of vertical parallel bars that shows the frequency distribution graphically.


Let's Create a frequency distribution of the School variable using an R built-in database named painters


```{r}
head(painters)
```

```{r}
dim(painters)
```


First Fetch the school column

```{r}
school <- painters$School
```


When we apply the table() function will compute the frequency distribution of the School variable 

```{r}
school_frequency <- table(school)
school_frequency
```



Now we apply the barplot function to produce its bar graph

```{r}
boxplot(school_frequency)
```




##### **Challenge**

Now we challenge ourselves, will create a bar graph of the total day calls in the customer signature dataset



```{r}
boxplot(churn$total_day_calls)
```






#### **Histogram Code Example 3.3**

A histogram shows the frequency distribution of a quantitative variable. The area of each bar is equal to the frequency of items found in each class.


Will Create a histogram using the faithful dataset 

```{r}
head(faithful)
```



Then applying the hist() function to produce the histogram of the eruptions variable 

```{r}
hist(faithful$eruptions)
```






Now let's Create a histogram of the total day minutes in the customer signature dataset

```{r}
hist(churn$total_day_minutes)
```





Let's have more fun with the churn data set
```{r}
hist(churn$total_eve_minutes)
hist(churn$total_night_minutes)
hist(churn$total_intl_minutes)
hist(churn$number_customer_service_calls)
hist(churn$total_eve_charge)
hist(churn$total_intl_charge)
hist(churn$total_night_charge)
```


