---
title: "DBSCAN Clustering"
author: "Dynasty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **DBSCAN Clustering**

Performing DBSCAN Clustering on the given IRIS Data set.

loading library necessary
```{r}
library("dbscan")
```

Loading the data set
```{r}
library(data.table)
m<-read.csv("http://bit.ly/IrisDataset")
head(m)
```


Removing class labels
```{r}
m1<-m[,c(1,2,3,4)]
head(m1)
```

 Applying our DBSCAN algorithm. We want minimum 4 points with in a distance of eps(0.4)
```{r}
db<-dbscan(m1,eps=0.4,MinPts = 4)
```

Printing out the clustering results
```{r}
print(db)
```


Plotting our clusters
```{r}
hullplot(m1,db$cluster)
```


## **Challenge 1**


Loading the data set
```{r}
library(data.table)
df1 <- fread("http://bit.ly/MSDBSCANClusteringDataset")
head(df1)
```
```{r}
anyNA(df1)
```
a. Removing class labels
```{r}
df11<-df1[,c(1,2,3,4,5,6,7)]
head(df11)
```

b. Applying our DBSCAN algorithm. We want minimum 4 points with in a distance of eps(0.4)
```{r}
db<-dbscan(df11,eps=0.4,MinPts = 4)
```

c. Printing out the clustering results
```{r}
print(db)
```

d. Plotting our clusters
```{r}
hullplot(df11,db$cluster)
```

## **Challenge 2**

Loading the data set
```{r}
library(data.table)
df2 <- fread("http://bit.ly/MSDBSCANClusteringDataset2")
head(df2)
```

```{r}
anyNA(df2)
```

a. Removing class labels
```{r}
df21<-df2[,c(2,3,4,5)]
head(df21)
```


b. Applying our DBSCAN algorithm. We want minimum 4 points with in a distance of eps(0.4)
```{r}
db<-dbscan(df21,eps=0.4,MinPts = 4)
```

c. Printing out the clustering results
```{r}
print(db)
```

d. Plotting our clusters
```{r}
hullplot(df21,db$cluster)
```

## **Challenge 3**

Loading the data set
```{r}
library(data.table)
df3 <- fread("http://bit.ly/MSDBSCANClusteringDataset3")
head(df3)
```

```{r}
anyNA(df3)
```

a. Applying our DBSCAN algorithm. We want minimum 4 points with in a distance of eps(0.4)
```{r}
db<-dbscan(df3,eps=0.4,MinPts = 4)
```

b. Printing out the clustering results
```{r}
print(db)
```

c. Plotting our clusters
```{r}
hullplot(df3,db$cluster)
```

