---
title: "Hierarchical Clustering"
author: "Dynasty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Hierarchical Clustering**
 
## Example 

Implement the hierarchical clustering algorithm using the Arrests data set

```{r}
data("USArrests")
```

Remove any missing value (i.e, NA values for not available) That might be present in the data

```{r}
df <- na.omit(USArrests)
```

```{r}
head(df)
```


Before hierarchical clustering, we can compute some descriptive statistics

```{r}
desc_stats <- data.frame(
  Min = apply(df, 2, min),    # minimum
  Med = apply(df, 2, median), # median
  Mean = apply(df, 2, mean),  # mean
  SD = apply(df, 2, sd),      # Standard deviation
  Max = apply(df, 2, max)     # Maximum
)
desc_stats <- round(desc_stats, 1)
head(desc_stats)
```

We note that the variables have a large different means and variances. 
This is explained by the fact that the variables are measured in different units; Murder, Rape, and Assault are measured as the number of occurrences per 100 000 people, and UrbanPop is the percentage of the state’s population that lives in an urban area.
They must be standardized (i.e., scaled) to make them comparable. Recall that, standardization consists of transforming the variables such that they have mean zero and standard deviation one.

As we don’t want the hierarchical clustering result to depend to an arbitrary variable unit, we start by scaling the data using the R function scale() as follows

```{r}
df <- scale(df)
head(df)
```

First we use the dist() function to compute the Euclidean distance between observations, d will be the first argument in the hclust() function dissimilarity matrix

```{r}
d <- dist(df, method = "euclidean")
```


We then hierarchical clustering using the Ward's method

```{r}
res.hc <- hclust(d, method = "ward.D2" )
```


Lastly, we plot the obtained dendrogram
```{r}
plot(res.hc, cex = 0.6, hang = -1)
```






## **Challenge 1**

Using the US Arrests data sets in the above example, compute hierarchical clustering with other linkage methods, such as single, median, average, centro id, Ward’s and McQuitty’s.

a. single method
```{r}
res.hc <- hclust(d, method = "single" )
```


```{r}
plot(res.hc, cex = 0.6, hang = -1)
```

b. median method
```{r}
res.hc <- hclust(d, method = "median" )
plot(res.hc, cex = 0.6, hang = -1)
```

c. average method
```{r}
res.hc <- hclust(d, method = "average" )
plot(res.hc, cex = 0.6, hang = -1)
```

d. centroid method
```{r}
res.hc <- hclust(d, method = "centroid" )
plot(res.hc, cex = 0.6, hang = -1)
```

e. macquitty method
```{r}
res.hc <- hclust(d, method = "mcquitty" )
plot(res.hc, cex = 0.6, hang = -1)
```

## **Challenge 2**

Perform hierarchical clustering using the mtcars dataset
```{r}
df <- mtcars
head(df)
```
 
 Checking for nulls
```{r}
anyNA(df)
```
 
 
 Descriptive statistics
```{r}
desc_stats <- data.frame(
  Min = apply(df, 2, min),    # minimum
  Med = apply(df, 2, median), # median
  Mean = apply(df, 2, mean),  # mean
  SD = apply(df, 2, sd),      # Standard deviation
  Max = apply(df, 2, max)     # Maximum
)
desc_stats <- round(desc_stats, 1)
head(desc_stats)

```
 
 
 
 Scaling
```{r}
df <- scale(df)
head(df)
```
 

```{r}
d <- dist(df, method = "euclidean")
res.hc <- hclust(d, method = "ward.D2" )
plot(res.hc, cex = 0.6, hang = -1)
```






```{r}
res.hc <- hclust(d, method = "single" )
plot(res.hc, cex = 0.6, hang = -1)
```

```{r}
res.hc <- hclust(d, method = "average" )
plot(res.hc, cex = 0.6, hang = -1)
```


```{r}
res.hc <- hclust(d, method = "centroid" )
plot(res.hc, cex = 0.6, hang = -1)
```

```{r}
res.hc <- hclust(d, method = "mcquitty" )
plot(res.hc, cex = 0.6, hang = -1)
```

## **Challenge 3**

Perform hierarchical clustering using the iris dataset
```{r}
df <- iris
head(df)
```

```{r}
anyNA(df)
```
```{r}
df$Species <- unclass(df$Species)
head(df)
```


```{r}
desc_stats <- data.frame(
  Min = apply(df, 2, min),    # minimum
  Med = apply(df, 2, median), # median
  Mean = apply(df, 2, mean),  # mean
  SD = apply(df, 2, sd),      # Standard deviation
  Max = apply(df, 2, max)     # Maximum
)

head(desc_stats)
```

```{r}
head(df)
```


```{r}
d <- dist(df, method = "euclidean")
res.hc <- hclust(d, method = "ward.D2" )
plot(res.hc, cex = 0.6, hang = -1)
```


```{r}
res.hc <- hclust(d, method = "single" )
plot(res.hc, cex = 0.6, hang = -1)
```

```{r}
res.hc <- hclust(d, method = "average" )
plot(res.hc, cex = 0.6, hang = -1)
```

```{r}
res.hc <- hclust(d, method = "centroid" )
plot(res.hc, cex = 0.6, hang = -1)
```


```{r}
res.hc <- hclust(d, method = "mcquitty" )
plot(res.hc, cex = 0.6, hang = -1)
```


## **Challenge 4**

Perform hierarchical cluster analysis on the given protein consumption.
Dataset url = http://bit.ly/HierarchicalClusteringDataset

```{r}
library(data.table)
data <- fread('http://bit.ly/HierarchicalClusteringDataset')
head(data)
```


```{r}
data$Country<- NULL
head(data)
```


```{r}
desc_stats <- data.frame(
  Min = apply(data, 2, min),    # minimum
  Med = apply(data, 2, median), # median
  Mean = apply(data, 2, mean),  # mean
  SD = apply(data, 2, sd),      # Standard deviation
  Max = apply(data, 2, max)     # Maximum
)
desc_stats <- round(desc_stats, 1)
head(desc_stats)
```


```{r}
data <- scale(data)
head(df)
```


```{r}
d <- dist(data, method = "euclidean")
res.hc <- hclust(d, method = "ward.D2" )
plot(res.hc, cex = 0.6, hang = -1)
```


```{r}
res.hc <- hclust(d, method = "single" )
plot(res.hc, cex = 0.6, hang = -1)
```

```{r}
res.hc <- hclust(d, method = "average" )
plot(res.hc, cex = 0.6, hang = -1)
```


```{r}
res.hc <- hclust(d, method = "centroid" )
plot(res.hc, cex = 0.6, hang = -1)
```

```{r}
res.hc <- hclust(d, method = "mcquitty" )
plot(res.hc, cex = 0.6, hang = -1)
```

