---
title: "R Basics - List and Vectors"
author: "Dynasty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lists and Vectors

## 1. Creating

### Lists

#### 1.1 Lists Code Example


###### List are R objects that contain elements of different types like numbers, strings, vectors, matrices, functions and lists too.


##### Will create a list using list() function
```{r}
alist <- list ("Red", "Blue", c(42,36,01), FALSE, 73.91, 128.6)
```




##### if we print our list

```{r}
alist
```


##### we get the enclosed list




## Vectors





###### will create three vectors; a, b, c

```{r}
a <- c(1,2,5.3,6,-2,4)                   # numeric vector 
b <- c("one","two","three")              # character vector 
c <- c(TRUE,TRUE,TRUE,FALSE,TRUE,FALSE)  # logical vector
```

##### Print out these vectors
```{r}
a
```
```{r}
b
```
```{r}
c
```






## 2. Naming

### Lists

#### 2.1 Naming Lists Code Example




##### The elements in a list can be given names which would allow one to have access to those elements as shown below

##### first we create list containing vector, matrix and a list

```{r}
list_data <- list(c("Jan","Feb","Mar"), matrix(c(3,9,5,1,-2,8), nrow = 2), list("green",12.3))
```

##### Next we give names to the elements
```{r}
names(list_data) <- c("1st Quarter", "A_Matrix", "A Inner list") 
```

##### Now we print out the list data below and you can see the output
```{r}
list_data
```



## Vectors

##### 2.2 Naming Vectors Code Example




##### You can give a name to the elements of a vector with the names() function as shown below

```{r}
a <- c("Serena Williams", "Tennis Player") 
names(a) <- c("Name", "Profession")
```

##### we can print out theses vectors and view the output as below

```{r}
a

```


##### Let's Have some more fun with this

```{r}
f <- c("Chimamanda Ngozi", "Non Fiction Author", "Americanah")
names(f) <- c("Name", "Genre", "Book")
```


##### We print this and view the output

```{r}
f
```



# 3. Selection

### Lists

#### 3.1 List Selection Code Example





##### Every element of the list can be accessed by the use of square brackets,and by numeric indices or by the logical index or by using element names.:


##### lets create a list

```{r}
ls <- list( first = 2, second = 4, third = list( fourth = 3.2, fifth = 6.4 ) )
```
##### Printing first and second elements of the list
```{r}
 ls [1:2]
```

##### IF you want the third element from the last inthe list

```{r}
ls[-3]
```

##### Another way to ask for first and second element on the list
```{r}
ls [c ("first", "second")]
```




## Vectors

#### 3.1 Vector Selection Code Example




##### First let's create a numeric vector

```{r}
a <- c(1,2,5.3,6,-2,4) 
```

##### IF we want the second and third element inthe vector
```{r}
a[c(2,3)]
```



# 4. Adding

### Lists

#### 4.1 Adding Lists Code Example




##### we can add an element only at the end of a list

##### First lets create a vector and a matrix

```{r}
list_data <- list(c("Jan","Feb","Mar"), matrix(c(3,9,5,1,-2,8), nrow = 2))
```

##### now we add an element

```{r}
list_data[4] <- "New element" 
list_data[4]
```





### Let's  challenge ourselves

##### First we create an empty list 
```{r}
months_of_the_years <- c()
```

##### Then we add the twelve months of the year
```{r}
months_of_the_years[1] <- "January"
```
```{r}
months_of_the_years
```
```{r}
months_of_the_years[2] <- "February"
months_of_the_years[3] <- "March"
months_of_the_years[4] <- "April"
months_of_the_years[5] <- "May"
months_of_the_years[6] <- "June"
months_of_the_years[7] <- "July"
months_of_the_years[8] <- "August"
months_of_the_years[9] <- "September"
months_of_the_years[10] <- "October"
months_of_the_years[11] <- "November"
months_of_the_years[12] <- "December"
```
#### Let's see our full list
```{r}
months_of_the_years
```

### Vectors

#### 4.2 Adding Vectors Code Example

#### Vectors can be combined via fuction c

# first let's create our vectors
```{r}
p = c(1, 2, 3)
q = c("aa", "bb", "cc")
```

#### now lets combine the vectors
```{r}
c(p, q)
```

##### let's put that into practice

### First we create our vectors
```{r}
a <- c("Serena Williams", "Tennis Player") 
names(a) <- c("Name", "Profession")
```

##### combining the two
```{r}
c(a, a)
```

