---
title: "Matrices"
author: "Dynasty"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Matrices

A matrix is a two dimensional data structure that is similar to a vector but additionally contains the dimension attribute.

## 1. Creating

When creating a matrix, the last two arguments to matrix tell it the number of rows and columns the matrix should have.


#### 1.1 Matrix Code Example

You can use the byrow=TRUE argument to create a matrix by rows instead of by columns as shown below:

##### Let's create a matrix mymat

```{r}
mymat <- matrix(1:12,4,3)
```

##### printing out matrix
```{r}
mymat
```


#### 1.2 Matrix Code Example


##### Let's use the byrow=TRUE argument to create a matrix by rows instead of by columns as shown below

```{r}
mymat <- matrix(1:12,ncol=3,byrow=TRUE)
```


##### print our matrix
```{r}
mymat
```


## 2. Naming

#### 2.1 Matrix Naming Code Example

#### # In order to remember what is stored in a matrix, you can add the names of the columns and rows. This will also help you to read the data as well as select elements from the matrix.

#### Let's create our matrix
```{r}
kenya <- c(460.998, 314.4) 
ethiopia <- c(290.475, 247.900) 
chad <- c(309.306, 165.8)
```

#### Then create a matrix geography matrix
```{r}
geography_matrix <- matrix(c(kenya, ethiopia, chad), nrow = 3, byrow = TRUE)
```

##### Now let's start naming
```{r}
location <- c("Lat", "Long")
countries <- c("Kenya", "Ethiopia", "Chad")
```

##### Now let's name the columns with location
```{r}
colnames(geography_matrix) <- location
```

### Let's name the rows with countries
```{r}
rownames(geography_matrix) <- countries
```

##### Now let's print final product
```{r}
geography_matrix
```

#### Now let's create a matrix family with column names Name, Age, Gender and Occupation.
```{r}
Matriachy <- c("Matriachy", 60, "F", "Queen")
Ryan <- c("Ryan", 7, "M", "Student")
Dynasty <- c("Dynasty", 30, "F", "Scientist")
Migan <- c("Migan", 26, "F", "Chef")
Wolfcode <- c("Wolfcode", 19, "M", "Engineer")
```
```{r}
family <- matrix(c(Matriachy, Ryan, Dynasty, Migan, Wolfcode), nrow = 5, byrow = TRUE)
```
```{r}
Header <- c("Name", "Age", "Gender", "Occupation")
```
```{r}
Name <- c("Matriachy", "Ryan", "Dynasty", "Migan", "Wolfcode")
```

```{r}
colnames(family) <- Header
```
```{r}
rownames(family) <- Name
```

```{r}
family
```


## 3. Adding a Column

You can add a row to a matrix using the rbind() function.

#### 3.1 Adding a Column Code Example

#### let's create an x matrix
```{r}
x <- matrix(1:9, nrow = 3)
```

#### Then ass a column
```{r}
```


```{r}
cbind(x, c(1, 2, 3))
```



##### Let's Add a column residence to your fictional family matrix 

```{r}
cbind(family, c("Arabia Terra", "Amazonis Planitia", "Amazonis Planitia", "Arabia Terra", "Amazonis Planitia"))
```

## 4. Adding a Row

#### 4.1 Adding a Row Code Example

#### Let's create a matrix
```{r}
x <- matrix(1:9, nrow = 3)
```

#### now let's add a row using cbind
```{r}
rbind(x,c(1,2,3))

```
#### Let's Add a fictional character to our fictional family matrix
```{r}
rbind(family, c("Ruby", 3, "F", "Youngling", "Amazonis Planitia"))
```

## 5. Selecting a Matrix

To select an element of a matrix, one needs to specify both the row and the column as shown:

#### Let's first select a matrix
```{r}
x <- matrix(1:9, nrow = 3)
```
#### let's Select the elements from the above matrix 

##### selecting the elemnt at 1st row 3rd column
```{r}
x[1,3]
```

#### Selecting second row
```{r}
x[2, ]
```

#### selecting 3rd column
```{r}
x[ ,3]
```

#### lets now select the last member of the family
```{r}
family[5, ]
```
##### selecting the first member of the family
```{r}
family[1, ]
```

## 6. Operations


### Matrix addition &amp; subtraction

#### 6.1 Matrix Addition &amp; Subtraction Code Example

 Matrix addition and subtract require the matrices to have the same dimensions.
 
##### Let's create matrices x and y
 
```{r}
x <- matrix(c(3, 9, -1, 4, 2, 6), nrow = 2)
y <- matrix(c(5, 2, 0, 9, 3, 4), nrow = 2)
```

### Print x matrix
```{r}
x
```

##### printing y matrix
```{r}
y
```


#### Now we Add the matrices and print out the result
```{r}
x + y
```

#### we even subtract the matrics
```{r}
x - y
```



### Matrix Multiplication &amp; Division

#### 6.2 Matrix Multiplication &amp; Division Code Example

Let's create two 2 x 3 matrices x and y

```{r}
x <- matrix(c(3, 9, -1, 4, 2, 6), nrow = 2)
y <- matrix(c(5, 2, 0, 9, 3, 4), nrow = 2)
```

#### Multiply the matrices and assign it to variable z
```{r}
z <- x * y
```

#### printing z
```{r}
z
```


#### Let also divide the matrices
```{r}
z <- x/y
```
```{r}
z
```

