---
title: "tSNE"
author: "Dynasty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **# t-Distributed Stochastic Neighbor Embedding (t-SNE)**

Will  apply the t-SNE to the MNIST dataset.

Loading the data set.
```{r}
train<- read.csv('http://bit.ly/MNISTDataset') 
```

Previewing the data set
```{r}
head(train)
```



```{r}
library(Rtsne)
```

Curating the database for analysis 
```{r}
Labels<-train$label
train$label<-as.factor(train$label)
```

Plotting
```{r}
colors = rainbow(length(unique(train$label)))
names(colors) = unique(train$label)
```

Executing the algorithm on curated data
```{r}
tsne <- Rtsne(train[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500)
```

Getting the duration of execution
```{r}
exeTimeTsne <- system.time(Rtsne(train[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500))
```

Plotting our graph and closely examining the graph
```{r}
plot(tsne$Y, t='n', main="tsne")
```


```{r}
plot(tsne$Y, labels=train$label, col=colors[train$label])
```




## **Challenge 1**
Apply t-SNE on the Iris data set. This help us to decide whether classes are separable in some linear or nonlinear representation.

Loading the data set
```{r}
Iris <- iris
head(Iris)
```


```{r}
Labels<-Iris$species

Iris$species<-as.factor(Iris$species)
```


```{r}
colors = rainbow(length(unique(Iris$species)))
names(colors) = unique(Iris$species)
```

```{r}
tsne <- Rtsne(Iris[,-4], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500)
```


```{r}
```


```{r}
```

```{r}
```

## **Challenge 2**
Apply t-SNE to the digits data set to visualize clusters of handwritten digits.

Loading the data set
```{r}
traindata <- fread("http://bit.ly/DigitsDataset", sep=",")
head(traindata)
```


```{r}
colors = rainbow(length(unique(traindata$class)))
names(colors) = unique(traindata$class)
```


```{r}
tsne <- Rtsne(traindata[,-65], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500)
```


```{r}
exeTimeTsne <- system.time(Rtsne(train[,-65], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500))
```


```{r}
plot(tsne$Y, t='n', main="tsne")
plot(tsne$Y, labels=traindata$class, col=colors[traindata$class])
```



