---
title: "Duplicates"
author: "Dynasty"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Duplicated Data

### 1. Identifying Duplicated Data

R checks for duplicates across rows through the duplicated() function. 

#### Identifying Duplicated Data Code Example 1.1 


Let's identify dupliccates inthe data below

```{r}
x1 <- c(2, 4, 5, 6)
x2 <- c(2, 3, 5, 6)
x3 <- c(2, 4, 5, 6)
x4 <- c(2, 4, 5, 6)
df <- data.frame(rbind(x1, x2, x3, x4))
df
```



Let's now find duplicated rows in our dataset

```{r}
duplicated_rows <- df[duplicated(df),]
```



we print out the variable to see if we have any duplicated rows

```{r}
duplicated_rows 
```


We have two duplicated rows


Removing these duplicated rows in the dataset or showing these unique items and assigning to a variable unique_items below

```{r}
unique_items <- df[!duplicated(df), ]
unique_items
```



 we can also remove duplicated rows in the dataset or show the unique items using the unique() function

```{r}
unique_items2 <- unique(df)
unique_items2

```



### Challenge 1

We are going to display and delete the only duplicate records in the iris dataset below:

```{r}
head(iris)
```

```{r}
dim(iris)
```


```{r}
iris_duplicated_rows <- iris[duplicated(iris),]
iris_duplicated_rows
```

We only have one duplicated row, so going to drop this duplicated row

```{r}
unique(iris)
```


Our row count has changed from 150 to 149 meaning the duplicated row has been dropped



### Challenge 2

Will check for duplicates and drop them in the following dataset

```{r}
library("data.table")
video_games <- fread('http://bit.ly/VideoGamesDataset')
```


Get a grimpse of our dataset

```{r}
head(video_games)
```


Number of rows and columns

```{r}
dim(video_games)
```


Now we check for duplicates
```{r}
Video_game_duplicated <- video_games[duplicated(video_games),]
Video_game_duplicated
```



We have 707 rows duplicated, we are going to drop them
```{r}
unique(video_games)
```

