---
title: "Nulls"
author: "Dynasty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```


### Missing Data

## 1. Finding Missing Values


In R, the missing values are shown by the symbol NA. To identify missing values in your dataset the function is.na() is normally used.

#### Finding Missing Values Example 1.1 
```{r}
#Install tidyr package
 
install.packages('tidyr')
 
 
#Load the library
 
library(tidyr)
```


Let's create a dataset dt

```{r}
Name <- c("John", "Tim", NA)
Sex <- c("men", "men", "women")
Age <- c(45, 53, NA)
dt <- data.frame(Name, Sex, Age)
```


We print it out to view

```{r}
dt
```

Now let's identify missing data in our dataset

```{r}
is.na(dt)
```

We have a TRUE meaning we have null values

We can also check the total number of missing values in each column with the code below

```{r}
colSums(is.na(dt))
```

So we have 1 for name column and one for age column but none for sex column


## 2. Dealing with Missing Values


Using na.omit() to omit all rows containing missing values.

#### Dealing with Missing Values Code Example 2.1

First let's see all rows with no missing values
```{r}
na.omit(dt)
```

We have two clean rows

Now let's record/fill the missing values in a column with a number
```{r}
dt$Age[is.na(dt$Age)] <- 99

dt
```

So we have filled the missing value inthe age column with 99


We can also try filling the missing value with mean value of the column, as shown below

```{r}
dt$Age[is.na(dt$Age)] <- mean(dt$Age, na.rm = TRUE)
```

Let's see the magic

```{r}
dt
```

We now have 49 which its the mean of 45 and 53

Using the given bus dataset below, will find the missing values of the payment_method and travel_to columns with athen appropriate values

First we import our data
```{r}
library("data.table")
bus_dataset <- fread('http://bit.ly/BusNairobiWesternTransport')
```
A sneak peek of how the data looks like

```{r}
head(bus_dataset)
```

Our focus is on two particular columns payment method and travel to columns. 


Lets just call the whole data frame


```{r}
is.na(bus_dataset)
```
We have no nulls in our dataframe.


Let,s see how we call a specific column

```{r}
is.na(bus_dataset$payment_method)
```
```{r}
is.null(bus_dataset$travel_to)

```


Let's have more fun by cleaning the dataset below
Dataset url = http://bit.ly/MS-PropertyDataset

```{r}
property <- fread('http://bit.ly/MS-PropertyDataset')
```
Let's see our dataset

```{r}
property
```

We now check for missing values

```{r}
is.na(property)
```
We see we have a few missing values in our dataset in column: PID, ST_NUM , NUM_BEDROOMS, NUM_BATH,SQ_FT and from the dataframe all our values inthese columns are numeric

For PID column there is a sequence so will add 100005000 just for flow

```{r}
property$PID[is.na(property$PID)] <- 100005000
```

Let's see if this is in effect

```{r}
property
```


Our efforts are paying off


We move to the next column ST_num and we see there is flow so will use fill() fucntion which fill the missing value with the previous value
```{r}
property%>%fill(ST_NUM)
```


Move on to the next column

